---
title: "HW1"
author: "Rory Michelen"
date: "September 5, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(tinytex)
Sys.setenv(PATH = paste(Sys.getenv("PATH"), "C:\\Users\\Rory\\AppData\\Local\\MiKTeX\\2.9\\miktex\\bin\\x64", sep=.Platform$path.sep))
```

# Question 1
## Part 1
Using the law of total probability, we can express the probability that system, S will be operational as a function of t. First, some definitions:

$S$ = Event that the system will be functional  
$E_{i}$ = Event that component $E_{i}$ will be functional  
$H_1= E_{5}(t)$  
$H_2=H_1^c$  

$P(S)=P(S \mid H_1)P(H_1)+P(S \mid H_2)P(H_2)$  

$P(S \mid H_1)=P(E_1)+P(E_3)-P(E_1)P(E_3)=e^{-t}+e^{-t/2}-e^{-3t/2}$

$P(H_1)=P(E_5)=e^{-t}$  

$P(S \mid H_2)=P(E_1)P(E_2)+P(E_3)P(E_4)-P(E_1)P(E_2)P(E_3)P(E_4)=e^{-3t}+e^{-5t/6}-e^{-23t/6}$  

$P(H_2)=1-P(E_5)=1-e^{-t}$  

$P(S)=(P(E_1)+P(E_3)-P(E_1)P(E_3))*P(E_5)+(P(E_1)P(E_2)+P(E_3)P(E_4)-P(E_1)P(E_2)P(E_3)P(E_4))*(1-P(E_5))$  
$P(s)=(e^{-t}+e^{-t/2}-e^{-3t/2})*e^{-t}+(e^{-3t}+e^{-5t/6}-e^{-23t/6})*(1-e^{-t})$  

Well, that was fun to write in LaTex! Let's move on to computation:


```{r Q1 Shared Code}
# Question 1
t<-seq(0,1,by=0.001)
l<-c(1,2,1/2,1/3,1)

prob_e<-function (t,l){
  return(exp(-1*l*t))
}

probs.at.t<-as.data.frame(t)
probs.at.t$H1<-prob_e(t,5)
probs.at.t$H2<-1-probs.at.t$H1
probs.at.t$s.gv.H1<-prob_e(t,1)+
                      prob_e(t,3)-
                      (prob_e(t,1)*prob_e(t,3))
probs.at.t$s.gv.H2<-(prob_e(t,1)*prob_e(t,2))+
                      (prob_e(t,3)*prob_e(t,4))-
                      (prob_e(t,1)*prob_e(t,2)*prob_e(t,3)*prob_e(t,4))

probs.at.t$s<-with((H1*s.gv.H1)+(H2*s.gv.H2),data=probs.at.t)
```

```{r Part A Plot}
probs.at.t%>%
  ggplot(aes(x=t,y=s))+
    geom_line(group=1)+
    labs(x="Time",
         y="Probability of Succes",
         title="Probability of Success over Time")

```

Below is the computation for probability of event S at time t=1/2.

```{r Part A time specific value,fig.show='hide'}
# Time =1/2
probs.at.t%>%
  dplyr::filter(t==0.5)%>%
  select(s)
```

Using this code, we get

$P(s)=0.28$ for t=1/2


## Part B

Using the world-famous bayes formula, we can derive the formula for the probability that component 5 was successful at time t=1/2

$P(E_5 \mid S)=\frac{P(S \mid H_1)}{P(S)}P(H_1)$

Each of these probabilities were expressed as a function of t in part A. Below is the computation for time t=1/2


```{r Part B}
# Part B
probs.at.t$H1.gv.s<-with(s.gv.H1*H1/s,data=probs.at.t)

probs.at.t%>%
  dplyr::filter(t==0.5)%>%
  select(H1.gv.s)
```

And we get the result of

$P(E_5 \mid S)=0.20$ for t=1/2

# Question 2

First, I'll define some important events

Let $B_{i,j}$ = Event that the $i^{th}$ item is selected from batch j  
Let $C_i$ = the event that the $i^{th}$ item is conforming  

Next, some probabilities that will be useful later:  
$P(B_{1,j})=0.5$ for $j \in \{1,2\}$ 

$P(C_1 \mid B_{1,1})=1.0$  

$P(C_1 \mid B_{1,2})=0.8$ 

$P(C_1)=P(C_1 \mid B_{1,1})P(B_{1,1})+P(C_1 \mid B_{1,2})P(B_{1,2})=(1*0.5)+(0.8*0.5)=0.9$  

Now, let's formulate the proposed question and plug in the above values. We were asked to find: 

$P(C_2^c \mid C_1)$  

$P(C_2^c \mid C_1)=1-P(C_2 \mid C_1)$   

$P(C_2 \mid C_1)=P(C_2 \mid C_1,B_{1,1})*P(B_{1,1}|C_1)+P(C_2 \mid C_1,B_{1,2})*P(B_{1,2}|C_1)$  

$P(C_2 \mid C_1,B_{1,1})=P(C_2 \mid B_{1,1})=1$  
$P(C_2 \mid C_1,B_{1,2})=P(C_2 \mid B_{1,2})=0.8$  

$P(B_{1,1}|C_1)=\frac{P(C_1\mid B_{1,1})}{P(C_1)}P(B_{1,1})=\frac{1}{0.9}0.5=\frac{5}{9}$  
$P(B_{1,2}|C_1)=\frac{P(C_1\mid B_{1,2})}{P(C_1)}P(B_{1,2})=\frac{0.8}{0.9}0.5=\frac{4}{9}$  

$P(C_2 \mid C_1)=1*\frac{5}{9}+\frac{8}{10}*\frac{4}{9}=\frac{41}{45}$

$P(C_2^c \mid C_1)=\frac{4}{45}=0.89$

# Question 3

Let $P_i$ denote the event that our algorithm predicted class i for $i \in \{0,1\}$  
Let $A_i$ denote the event that an item selected from the population is of class i for $i \in \{0,1\}$   
Let $S_i$ denote the even that an item selected from the sample is of class i for $i \in \{0,1\}$ 

It is given that:  
$P(A_0)=0.99$  

$P(A_1)=0.01$  

$P(S_0)=\frac{55}{120}$  

$P(S_1)=\frac{65}{120}$  

$P(P_1)=\frac{70}{120}$  

$P(P_1|S_1)=\frac{52}{70}$

We are asked to find $P(A_1 \mid P_1)$ 

$P(A_1 \mid P_1)=\frac{P(P_1 \mid A_1)}{P(P_1)}P(A_1)$  

$P(P_1 \mid A_1)=(P_1 \mid S_1)=\frac{52}{70}$  

$P(A_1 \mid P_1)=\frac{\frac{52}{70}}{\frac{70}{120}}*0.01=0.0135$







