---
title: "HW6"
author: "Rory Michelen"
date: "November 9, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(R2OpenBUGS)
library(knitr)
```

# Question 1
## Model Creation

To answer this question, I will model the lifetimes as a weibull distribution. The model is created in R2Open Bugs, below
```{r,message=FALSE}
# Read the input data
input.data<-read_csv('tongue.csv',col_names=FALSE)
colnames(input.data)<-c('dna.profile','duration','observed')
input.data<-input.data%>%
  mutate(lifetime=ifelse(observed==1,duration,NA),
         censored=ifelse(observed==0,duration,0))

lifetime<-input.data$lifetime
censored<-input.data$censored
profile<-input.data$dna.profile
N<-nrow(input.data)

# Specify the model
model<-function(){
  for(i in 1 : N){
    lifetime[i]~dweib(v,lambda[i])%_%C(censored[i],)
    lambda[i]<-exp(b0+b1*profile[i])
    median[i]<-pow(log(2)*exp(-b0-b1*profile[i]),1/v)
  }
  b0~dnorm(0,0.0001)
  b1~dnorm(0,0.0001)
  v~dexp(0.001)
}

model.file <- file.path(tempdir(), "model.txt") 
write.model(model, model.file)

data<-list("lifetime","censored","profile","N")
inits <- function() {list(b0=0,b1=0,v=1)}
params<-c("b0","b1","v","median")

out <- bugs(data, inits, params,model.file, n.iter=10000)
```


## Part A

The 95% Credible Set is provided below for all 3 parameters. For $b_1$ the credible set is [-1.368,0.370]

```{r}
out$summary%>%
  data.frame()%>%
  rownames_to_column('Parameter')%>%
  select(Parameter,mean,lower.bound='X2.5.',upper.bound='X97.5.')%>%
  dplyr::filter(Parameter %in% c('b0','b1','v'))%>%
  kable()
```


# Question 2

```{r}
X<-c(2, 1, 0, 2, NA, 3, 1, 0, 1, 2, 3, 0, 1, NA, NA)
Y<-c(NA, 16, 9, 17, 12, 22, 13, 8, NA, 19, 17, 11, 10, 20, 2)
N<-length(X)


model<-function(){
	for(i in 1:N){
		Y[i]~dpois(lambda[i])
		lambda[i]<-exp(beta[1]+beta[2]*X[i])
		X[i]~dpois(2)
	}
	for(i in 1:2){
		beta[i]~dnorm(0,0.0001)
	}
  mean.w.4routes<-exp(beta[1]+beta[2]*4)
  predict.w.4routes~dpois(mean.w.4routes)
}

model.file <- file.path(tempdir(), "model.txt") 
write.model(model, model.file)

data<-list("X","Y","N")
inits <- function() {list(beta=c(0,0),X=c(NA, NA, NA, NA, 2, NA, NA, NA, NA, NA, NA, NA, NA, 2, 2))}
params<-c("beta","X","Y","mean.w.4routes","predict.w.4routes")

out <- bugs(data, inits, params,model.file, n.iter=10000)


```
## Part A
The output of the above model is provided below. The deviance of our model is 109.588 with an 95% credible set of [105.5,118.0]
```{r}
out$summary%>%
  data.frame()%>%
  select(mean,lower.bound='X2.5.',upper.bound='X97.5.')%>%
  rownames_to_column("Parameter")%>%
  kable()
```

## Part B
The mean expected number of broken packages for a shipment with 4 routes has a credible interval of [18.72,40.79] with a mean of 28.29.

## Part C
The predicted number of broken packages for a shipment with 4 routes has a credible interval of [15,45] and a mean of 28.32.  

While the mean and the prediction are centered roughly at the same value, the credible set of the prediction is wider than that of the mean. This is because the prediction has an additional source of uncertainty. 

There is a shared source of uncertainty: What is the true value of the paramter lambda. However, the prediction has an additional source of uncertainty: given a known distribution with fixed lambda, what will the value of the next observation be?

## Part D

The table below provides estimates of all missing data

```{r}
out$summary%>%
  data.frame()%>%
  select(mean,lower.bound='X2.5.',upper.bound='X97.5.')%>%
  rownames_to_column("Parameter")%>%
  dplyr::filter(substr(Parameter,1,1)%in% c('X','Y'))%>%
  kable()
```







